\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{setspace}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{caption}
\captionsetup[figure]{font=footnotesize}
\usepackage{csvsimple}
\usepackage{amssymb}
\usepackage{textgreek}
\usepackage{amsmath}
\usepackage{harvard}
\usepackage{parskip}
\usepackage[a4paper,width=160mm,top=20mm,bottom=20mm,bindingoffset=6mm]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[CE]{Model Selection in Microbial Population Growth}

\linespread{1.25}

\begin{document}

\input{Titlepage}

\newpage

\section{Declaration}
All raw data collected were from the simulation I created for my project. The simulation requires the input of genetic model systems, and mathematical equations used to derive output. The model system and equations used in the simulation were sourced from the work of /textit{Omholt et al,} and \textit{Gjuvsland et al}. I was reponsible for data processing, cleaning and analysis. All analyses presented in the paper are from the simulations, with the help of my supervisor.

\section{Acknowledgements}
I would like to firstly thank Dr. Scott Rifkin for being a wonderful supervisor and guiding me throughout the project. This includes understanding background knowledge, results and overall research purposes. Secondly, thank you to Dr. Thomas Bell for agreeing to be my internal supervisor, making sure I am aware of the process of the project and ensuring my safety during such difficult times.
\\Finally I would like to thank the laboratory of Dr. Rifkin- Antonia Darragh, Jessica Bloom, Alexis Cugini, Yang Bing and Rachel Goodridge for being very welcoming and having wonderful and insightful weekly meetings. Good luck with everything!

\newpage

\section{Abstract}

\newpage

\section{Introduction}
\input{Sections/Introduction}

\section{Methods}
\input{Sections/Methods}

\section{Results}

\section{Discussion}

\section{Conclusion}

\newpage

\section{Data Code and Availability}
All code can be found in the GitHub Repository:\\
{https://github.com/matthewcampos/CMEECourseWork.git}

\newpage

\bibliographystyle{agsm}
\bibliography{references}

\newpage

\section{Appendix I}

\section{Appendix II}

\end{document}
